---
title: "Fancier Maps"
editor: visual
---

We can use leaflet to make some fancier maps.

```{r}
#install.packages(leaflet)
library(leaflet)
```

[Manual here](https://cran.r-project.org/web/packages/leaflet/leaflet.pdf) [Here is a text on leaflet.](https://bookdown.org/nicohahn/making_maps_with_r5/docs/leaflet.html)

```{r}
leaflet()|>
  addTiles()
```

Let's put Smith College on the map. [Here's the lat and long.](https://www.google.com/search?q=smith+college+latitude+and+longitude&client=firefox-b-1-lm&sca_esv=e89764eee12a19ad&ei=ibXQZuWTDJf8ptQP-6iyqQw&oq=smith+college+latitude+and+long&gs_lp=Egxnd3Mtd2l6LXNlcnAiH3NtaXRoIGNvbGxlZ2UgbGF0aXR1ZGUgYW5kIGxvbmcqAggAMggQIRigARjDBDIIECEYoAEYwwRI9ylQ_RhY2R5wA3gBkAEAmAGeAaABkAWqAQMwLjW4AQHIAQD4AQGYAgigAsUFwgIKEAAYsAMY1gQYR8ICChAhGKABGMMEGArCAgQQIRgKmAMAiAYBkAYIkgcDMy41oAfwFg&sclient=gws-wiz-serp)

```{r}
leaflet()|>
  addTiles() |>
  addMarkers(lng = 72.6403,
             lat = 42.3163,
             popup = "Home of the Pioneers",
             label = "Smith College")
```

Whooops. We need a negative sign to indicate west.

```{r}
leaflet()|>
  addTiles() |>
  addMarkers(lng = -72.6403,
             lat = 42.3163,
             popup = "Home of the Pioneers",
             label = "Smith College")
  
  


```

```{r}
leaflet()|>
  addTiles() |>
  addMarkers(lng = -72.6403,
             lat = 42.3163,
             popup = "Home of the Pioneers",
             label = "Smith College")
```

You try:

1.  Put your home city and Smith College on the same map.

2.  Use `awesomeIcons()` to make your home look like a home and smith look like a college.

    i.  You will first need to save the icon an object. (Maybe the home is called home).

        Hint: Use icon "education", "book", "sleeping" for the college icon.

    ii. Then you will need to use `addAwesomeMarkers()`.

3.  Finally add this webpage to your index.qmd and make sure it renders correctly.

```{r}
# My awesome icons are not working on the rendered map
# Created objects for Smith & home icons
education <- awesomeIcons(icon = "education",
                          iconColor = "black")
home <- awesomeIcons(iconColor = "white")

# Used awesome markers to mark it on the map
leaflet()|>
  addTiles() |>
  addAwesomeMarkers(lng = -72.6403,
             lat = 42.3163,
             icon = education,
             popup = "Home of the Pioneers",
             label = "Smith College") |>
  addAwesomeMarkers(lng = 123.4330,
             lat = 41.8048,
             icon = home,
             popup = "Original Hometown",
             label = "Shenyang")
```
