---
title: "SDS 192 Mini-Project 1"
subtitle: "SDS 192"
author: 
- name: "Emily"
  affiliation: Smith College
- name: "Maisha"
  affiliation: Smith College
- name: Author Three
  affiliation: Smith College
date: "`r format(Sys.Date(), '%B %e, %Y')`"
editor: 
  mode: source
output: 
  html_document:
    fig_width: 7
    fig_height: 6
    fig_caption: true
    theme: lumen
    df_print: paged
    code_folding: show
---

```{r setup, include=FALSE}
library(tidyverse)
library(rscorecard)
library(RColorBrewer)

sc_key("Pbe6JRZgbpxV4xnEdrZcKjqYbf04Jbe9ZTSbdCcN")
```

```{r import}
scorecard <- sc_init() |>
  sc_year(2018) |>                 #Note how we are looking at only 2018 data here!
  sc_filter(stabbr == "MA") |>     #Note how we are looking at only Massachusetts data here!
  #The line below shows variable selection (there are lots of variables)
  sc_select(unitid, city, control, ugds, ret_ft4) |>
  sc_get()
```

```{r recode}
#  Here's an example of how to recode the control variable

# We are renaming the column control_text from control
scorecard$control_text <-
  #The recode function does the work. It calls 1 "Public", 2 " Private nonprofit",etc. 
  recode(
    scorecard$control, 
    "1" = "Public", 
    "2" = "Private Non-profit", 
    "3" = "Private for-profit",
    .default = NA_character_
  )
```

```{r}
view(scorecard)
```

```{r}
#Get the largest 50 colleges in MA
sc_largest_50 <- scorecard |>
  arrange(desc(ugds)) |> #orders descending
  head(50) #selects first 50

view(sc_largest_50)
```

# Plot 1

```{r plot_1}
# Plot-1 : Frequency distribution of type of colleges among the 50 largest colleges in MA
ggplot(data = sc_largest_50,
       map = aes(x = reorder(city, city, length), # reorders the cities with the highest number of UG schools from the largest 50
                 fill = control_text)) +
  geom_bar(width = 0.75) +
  
  # Labeling the aesthetics
  labs(x = "Cities",
       y = "Number of Colleges",
       fill = NULL,
       title = "Governance Structure of Colleges in Massachussetts",
       subtitle = "Of the 50 colleges with the largest undergraduate student population in 2018",
       caption = "The 'City' variable with the cities with the 50 colleges having the  largest undergraduate student population has been mapped to x-axis (arranged in descending order) and the y-axis reflects the counts of the colleges.
       The fill color maps the 'Public/Private Non-profit/Private For-profit' variable with the type of governance structure.
       
       Source: College Scorecard by the U.S. Department of Education (https://collegescorecard.ed.gov/)") +
  
  #formatting scales 
  scale_y_continuous(breaks = seq(0, 10, by = 1)) + #  1 college equal to 1 unit on y axis
 scale_fill_manual(values = c("#FF0000", "#0000FF")) + # Manually changes the control variable colors +
  #scale_fill_brewer(direction = -1, type = "qual", palette = "Set1")+
  coord_flip() + # Horizontal bar chart
  
  # Formatting the graph
  theme_minimal() +
  theme(
    # aligning the caption to the plot area instead of panel
    plot.caption.position = "plot", 
    
    # positioning the legends on the top
    legend.position = "top",
    
    # formatting texts and margins for the texts and titles
    axis.text = element_text(size = 6), 
    plot.title = element_text(face = "bold", 
                              hjust = "0.5"),  
    plot.margin = margin(t = 20, r = 50, b = 0, l = 40),
    axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)),
    axis.title.x = element_text(margin = margin(t = 20, r = 0, b = 0, l = 0)),
    plot.caption = element_text(color = alpha("black", 0.5),
                                face = "italic",
                                size = 7,
                                margin = margin(t = 20, r = -30, b = 20, l = 50)),
    
    # formatting grid color
    panel.grid = element_line(color = alpha("grey", 0.2)), 
  )
```


# Plot 2

```{r plot_2}
# Create plot 2 here

ggplot(data = sc_largest_50,
       aes(x = ugds,
           y = ret_ft4,
           color  = control_text)) +
  geom_jitter(size = 1.5, alpha = 0.75) + # Addressing Overplotting
  scale_y_continuous(labels = scales::percent)+
  labs(
    title = "Retention Rates in Comparison with the Undergraduate Student Number",
    subtitle = "Of the 50 colleges with the largest undergraduate student population in 2018",
       x = "Undergraduate Student Number",
       y = "Retention Rates (%)",
       color = NULL,
       caption = "The 'Number of Undergraduate Students' variable of the 50 colleges having the  largest undergraduate student population has been mapped to x-axis and the y-axis reflects the 'Retention Rate of Fulltime students' at 4yr colleges with the .
       The  color maps the 'Public/Private Non-profit/Private For-profit' variable with the type of governance structure.
       
       Source: College Scorecard by the U.S. Department of Education (https://collegescorecard.ed.gov/)") +
  scale_color_manual(values = c("#FF0000", "#0000FF")) + # Manually changes the control variable colors

 # Formatting the graph
  theme_minimal() +
  theme(
    # aligning the caption to the plot area instead of panel
    plot.caption.position = "plot", 
    
    # positioning the legends on the top
    legend.position = "top",
    
    
    # formatting texts and margins for the texts and titles
    axis.text = element_text(size = 6), 
    plot.title = element_text(face = "bold", 
                              hjust = "0.5"),  
    plot.margin = margin(t = 20, r = 50, b = 0, l = 40),
    axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)),
    axis.title.x = element_text(margin = margin(t = 20, r = 0, b = 0, l = 0)),
    plot.caption = element_text(color = alpha("black", 0.5),
                                face = "italic",
                                size = 7,
                                margin = margin(t = 20, r = -30, b = 20, l = 50)),
    
    # formatting grid color
    panel.grid = element_line(color = alpha("grey", 0.2))
  )

```

# Blog post

<i> “Does the governance structure of the largest colleges in Massachusetts affect the retention rate?” </i>

We investigated this question using the College Scorecard dataset. This data aims to increase transparency and help prospective students and families compare colleges based on their goals and requirements. The College Scorecard was produced by the U.S. Department of Education taking data from the federal reporting from institutions, data on federal financial aid, and tax information. These were taken directly or derived from the data reported to the Integrated Postsecondary Education Data System (IPEDS).

From the dataset, we will be investigating our question using the data of 50 colleges in Massachusetts with the largest undergraduate student population. Delving deeper, we will be exploring the governance structure and retention rates.
Firstly, we will group our selection of colleges based on the cities these are located in and then visualize the distribution of the type of governance structures – whether the colleges are Public, Private Non-profit, or Private For-profit.
Then, we will examine the correlation between retention rates and the undergraduate population based on the type of institution to help us draw conclusions from the data.

The first plot explores the governance structure of the largest 50 colleges in different cities of Massachusetts. The largest 50 schools are grouped with the cities these are located in and the length of the bars indicates the number of Public, Private Non-Profit, and Private For-Profit universities in each of those cities. Notably, there are no Private For-profit college in our selection, hence the bar chart omits the category. The plot concludes Boston has the highest number of big schools, followed by Worcester, Springfield, and Cambridge. The colors of the bars indicate that overall the number of public schools is higher than the number of private colleges in our selection of the largest MA colleges.

The second plot explores the correlation between the undergraduate population and the retention rate. For public colleges, we can see an inconclusive positive correlation between a larger undergraduate population and a higher retention rate as the points are scattered, showing that larger public universities are likely to have a higher retention rate. However, the points for private nonprofit colleges are in a larger cluster, signifying that there isn’t a direct and obvious correlation between undergraduate student body size and retention rate.

Assumptions and commitments that informed this data set include assuming that the colleges participate in Title IV programs (or meet similar criteria to participating institutions), and that these colleges report wholly correct data to the IPEDS. However, a lot of the data is missing as colleges did not report some data to the IPEDS. The educational institutions and organizations collecting the data have had a say regarding this dataset, and it is hard to find reliable data if the institutions don’t report their information. The benefits and harms of this dataset may be (shouldn’t be, but) that institutions can pick and choose the data they present, so that either data is missing, or it is positively skewed towards certain institutions when it’s beneficial. 




